Accesible, realistic genome simulation with selection using \stdpopsim

### 2025-03-12 analysis2_manuscript.tex
added discussion stuff

\newcommand{\tskit}{\texttt{tskit}\xspace}
\newcommand{\slim}{\texttt{SLiM}\xspace}
\newcommand{\msprime}{\texttt{msprime}\xspace}

@@

    % current numbers of DFEs / species / etc
    % ADK: check version number
    At this release (version 0.3XX) the \stdpopsim catalog comprises 27 species (i.e. genome representations)
    for which we have included 35 demographic models, 38 genetic maps, 12 DFEs, and 8 annotation objects. %ADK: checked this on 3/10/25
    Additions to the catalog are ongoing---we point the reader to our previously published 
    report detailing this effort \citep{lauterbur2023expanding}. We welcome contributions from the
    community and encourage those interested in contributing to visit https://popsim-consortium.github.io/stdpopsim-docs
    or contact one of the PopSim Consortium members for guidance.

@@

    To do this we simulated human genomes with and without natural selection.
    In both scenarios, we ran three replicate simulations
    under the out of Africa (OOA) demographic model of
    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
    Simulations with selection were implemented by modeling purifying selection on exons
    % IG: moved details to methods
    %, defined from the HAVANA group release 104 for the human genome,
    using a DFE inferred by \cite{kim2017inference}.
    A basic summary of nucleotide diversity from these simulations is shown in Figure \ref{fig:homsap-summary-stats},
    which shows that the presence of selection decreases polymorphism in simulations with selection only slightly
    at the genome-wide level, but more substantially within exonic regions under both a constant population size
    and the out-of-Africa demographic model.
    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
    % IG: I commented this out because it appeared distracting. We can add a note in the end of this analysis
    % However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.

    Although the estimates produced by the four methods differ from one another, 
    for each method the estimates produced from data with selection (Figure
    \ref{fig:human-demography}; orange) and without selection (Figure
    \ref{fig:human-demography}; blue) are fairly similar.
    Thus, this particular parameterization of background selection
    in combination with selection only at exonic sites in the out-of-Africa demographic model does not
    appear to notably bias the estimates produced by these methods.
    As noted in previous studies that compared demography inference methods
    on neutral simulations \citep{adrion2020community},
    we see that each method has its strengths and weaknesses.
    Methods based on the sequential Markovian coalescent (\smcpp and \msmc)
    appear to overal produce the most accurate trajectories of $N_e(t)$
    in this setting, although they tend to under-estimate the population
    bottleneck associated with the population split.
    On the other hand, \stairway, which utilizes the site frequency spectrum,
    produces more noisy estimates,
    but it does not appear to under-estimate the population bottleneck.
    The inference of \gone, which utilizes identity by descent,
    is targeted toward demographic changes in more recent time frames ($\sim 200$ generations),
    explaining its noisy estimates for all but the most recent time periods.

@@

    In comparing demography inference methods, we see that the site-frequency-based method (\stairway) produces more accurate
    inference than the two methods based on the sequential Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
    Reagrdless of these differences, we see that background selection again does not
    appear to considerably influence the inference of any of the four methods for this species,
    which has a very different evolutionary history than humans.

@@

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
        \drscomment{I think it would be good to increase the font size for the A,B,C panel legends and also
        draw a horizontal line above panel C so it is clear that all three plots at the bottom are part of C.}
        \caption{Inferred versus simulated DFEs from simulated human genomes.
        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
        acting on exons (see \textbf{Methods}).
        DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
        by three different methods: \grapes, \polydfe , and \dadicli.
        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
        shown for each DFE inferred from all three simulated datasets,
        with median values marked by horizontal bars
        and simulated values represented by dashed horizontal lines.
        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
        white bars represent the distribution used in the simulation.
        }
        \label{fig:homsap-dfe.ooa}
    \end{figure}

@@

    \begin{figure}[b!]
        \centering
        \includegraphics[width=0.8 \textwidth]{figures/sweeps/relationship_power_cM.pdf}
        \caption{
        Power to detect selective sweeps 
        %IG: shortened the title line
        %at 100 locations along human
        %chromosome 1 under a three population out-of-Africa model
        %\citep{gutenkunst2009inferring}  when using null models with and
        %without purifying selection
        as a function of local recombination rate.
        This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
        but with the 100 genomic segments sorted according to their
        average recombination rates instead of position along chromosome 1.
        \igcomment{Specify how the curves were computed}
        %IG: copoed from the caption of fig 5
	Genomic segments were simulated with sweeps under a three population out-of-Africa model
        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
        Three methods for detecting sweeps were applied to simulated data:
        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
        2) \diploshic \citep{kern2018diplos} (middle row),
        and 3) reduced diversity ($\pi$) (bottom row).
        Power (true positive rate) is shown for these methods for the CEU and YRI
        samples (left and right respectively).
        The thresholds of the test statistics were set to control for
        $5\%$ false positive rate under a neutral null model (blue)
        and a null model with background selection from deleterious mutations in exons (red).
        }
        \label{fig:power-recomb}
    \end{figure}

@@

    Inference of the DFE was done here using exon annotations of the genome in question.
    A key distinction between our analyses and those done in practice
    is that we have complete knowledge of which sites are under selection in the simulation,
    thus our inferences represent a best case scenario. Empirical inference of the DFE is a much
    harder problem as some mutations assumed to be free of selection
    may in reality be under selection

    One important finding we made using the new stdpopsim selection API is that inference of the DFE is sensitive to the dominance of deleterious mutations.
    All three of the inference methods of the DFE applied here assume that deleterious mutations act in an additive fashion (h=0.5) by default.
    However, the simulated DFE for the vaquita genome included highly recessive mutations.
    Specifically, these simulations used a DFE where mutations with $s>0.01$ are quite recessive, 
    with $h<=0.01$, and where mutations with $0.001<s<0.01$ have $h=0.1$.
    We found that the three DFE inference methods underestimated $E[s]$ and overestimated the shape parameter (Figure \ref{fig:vaquita-dfe}).
    This performance is not surprising as decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$,
    or an apparent decrease in the amount of selection.
    Fitting a mis-specified DFE that assumes $h=0.5$ will then lead to underestimates of $E[s]$.
    Indeed, the distribution of $hs$ is much more accurately estimated (Figure \ref{fig:vaquita-dfe}b).
    Overall, these results are consistent with prior work showing that it is challenging to separately
    infer the DFE of $s$ along with $h$ \citep{veeramah2014evidence, kyriazis2024constraining, balick2022overcoming}
    and that a large proportion ($>5\%$) of recessive deleterious mutations can confound inferences of $s$ \citep{wade2023quantifying}.
    Given the plethora of evidence for recessive deleterious mutations 
    \citep{mukai1972mutation, agrawal2011inferences, huber2018gene, di2024revisiting,},
    these results suggest the need for care in DFE inference.
    Kyriazis and Lohmueller \citeyear{kyriazis2024constraining} inferred the DFE of $s$ for nonsynonymous mutations
    in humans when considering different values of $h$.
    Intriguingly, additive models showed a satisfactory fit to the SFS of nonsynonymous variants in humans
    and more complex models of dominance did not substantially improve the fit.
    Nevertheless, models including recessive deleterious mutations, like those simulated here for the vaquita genome,
    show a similar fit to the SFS and may be more biologically realistic than fully additive models,
    given the evidence for the presence of recessive deleterious mutations.
    More broadly, new computational approaches are needed to precisely quantify the DFE and the covariance of $h$ and $s$.

    Regarding sweep detection, the inclusion of selection models into \stdpopsim allowed us to readily compare the
    effectiveness of different methods on data reflecting the demographic history, recombination rate variation,
    and genomic locations of negatively selected sites present in human populations. These results clearly demonstrate
    the impact of recombination rate variation, and underscore the importance of training methods on data with variable
    recombination rates (e.g., following Schrider and Kern \citeyear{schrider2017soft}). In addition, our results
    illustrate a strength of likelihood methods like \sweepfinder, which do not require a predefined window size and
    thus will perform better than \diploshic---which only examines relative summarie of diversity across a window instead
    of the actual values of these diversity measures---on very strong sweeps that effect diversity across a wide stretch of
    chromosomes that extends far beyond \diploshic's window size.

@@

    \subsection*{Whole genome simulations}
    In this study we performed simulations of the human and vaquita porpoise genomes
    using the \stdpopsim{} framework. We ran replicate datasets for each genome/parameter combination,
    in which chromosomes were simulated separately from one another.
    All simulations were executed using the \slim engine with a burn-in of $10N$ generations
    (where $N$ is the initial population size), and a scaling factor of 1 (no scaling).
    All simulations were implemented using as part of a snakemake workflow,
    where the code utilized is similar to the one shown in Figure 1B.
    The complete simulation workflow is available at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/simulation.snake}
    (see \textbf{Data availablity} below).

    \subsubsection*{Simulations of complete human genomes}
    Simulation of human genomes used a genetic map from the HapMap Project
    \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
    the out of Africa (OOA) demographic model with archaic admixture from
    \cite{ragsdale2019models} (\stdpopsim label \texttt{OutOfAfricaArchaicAdmixture\_5R19}).
    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
    \cite{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
    (\cite{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
    Each simulation generated 100 diploid genomes from each of the three extant human populations (YRI, CEU, and CHB),
    with each genome consisting of 22 autosomes.
    Using these settings, we generated three replicate datasets without selection
    and three replicate datasets with selection.
    Three additional datasets, used in Figure \ref{fig:homsap-dfe.constant},
    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $10,000$.
    Each of these simulations generated 100 diploid genomes.
    The configuration file used in combination with our snakemake workflow
    for these simulations can be found at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_HomSap.yml}.

    \subsubsection*{Simulations of complete vaquita porpoise genomes}
    Simulation of vaquita porpoise genomes used 
    the two-epoch demographic model from \cite{robinson2022critically} (\stdpopsim label \texttt{Vaquita2Epoch\_1R22}).
    Since no genetic map is currently available for this species, we assumed a constant recombination rate of  $10^{-8}$
    across all chromosomes \citep{morin2021}.
    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
    \cite{robinson2022critically} (\stdpopsim label \texttt{Gamma\_R22}) to exons annotated from the vaquita genome assembly
    (\cite{morin2021}; \stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}).
    This DFE model implements a relationship between selection and dominance as follows:
    very deleterious mutations (with $s<-0.1$) were set to be recessive ($h=0$),
    mutations with $s\in [-0.1,-0.01)$ were set to have a dominance coefficient of $h=0.01$,
    mutations with $s\in [-0.01,-0.001)$ were set to have a dominance coefficient of $h=0.1$,
    and very mild mutations (with $s\in [-0.001,0)$) were set to have a nearly additive effect ($h=0.4$).
    This approach is similar to what was used in the simulations cunducted by \textcite{robinson2022critically}
    and is easily implemented in \stdpopsim's \texttt{DFE} class.
    Each simulation generate 100 diploid genomes,
    with each genome consisting of 21 autosomes.
    Using these settings, we generated three replicate datasets without selection
    and three replicate datasets with selection.
    Three additional datasets, used in Figure \ref{fig:vaquita-dfe.constant},
    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $3,500$.
    \igcomment{If we end up using Fig. S3, then we should add a line describing it's simulation here.}
    The configuration file used in combination with our snakemake workflow
    for these simulations can be found at 
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_PhoSin.yml}.

@@

    We assessed the accuracy of demographic inferences from three software packages:
    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent},
    and \smcpp \citep{terhorst2017robust}.
    For \msmc \citep{Schiffels2020}, we used a random sample of $N=6$ simulated genomes with 20
    iterations of the EM algorithm.
    For \gone \citep{santiago2020recent}, we set \texttt{max\_snps} to 500,000,
    the number of generations to 2,000, and the number of bins to 400, all representing default settings.
    For \stairway \citep{liu2020stairway} we did one run of theta estimation for each replicate dataset,
    and used a `dimFactor' of 5000, both of which represent default settings.
    For \smcpp \citep{terhorst2017robust} we used the unfolded SFS (`--unfold'), 
    along with the true mutation rate under which the simulations were generated. 
    The complete snakemake workflow for demographic inference can be found at
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/n_t.snake}.

@@

    % ADK: I've added a section below that describes the DFE inference methods used in this study.
    For comparison among methods, estimates were rescaled. 
    As described above, each DFE-inference method scales selection coefficients slightly differently.
    We thus rescaled each program's output to obtain a selection coefficient ($s$) consistent with the one used by \slim.
    The output of \grapes  was scaled by a factor of  $\mu / \theta$,
    and the output of \polydfe was saled by $2 \mu / \theta$, 
    where $\mu$ is the assumed mutation rate and 
    $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
    For \dadicli, we report $\alpha * \nu / (\theta / (4 \mu L p))$, where
    $\alpha$ and $\nu$ are the inferred shape and rate parameters of the \dadicli's inferred gamma distribution,
    $L$ is the total number of sites available for the analysis,
    $p$ is the proportion of sites under selection. 
    Complete details for the DFE inference analysis can be found in the
    snakemake workflow at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/dfe.snake}.

@@

    The three sweep-detection methods were associated with thresholds for their respective test statistics
    ($\pi$, \sweepfinder's CLR, or \diploshic's probability),
    which were determined using an empirical distribution representing a null model without sweeps.
    We considered two null models: a neutral evolution model (neutral),
    and a model with background selection from deleterious mutations in exons (BGS).
    Each null model was represented by a different set of $100,000$ simulations generated according to the appropriate model
    (see \textbf{Simulations with selective sweeps} above).
    Thus, for each test statistic we determined two thresholds corresponding to these two different null models (neutral and BGS)
    using the extreme $5\%$ quantile in each empirical distribution.
    Simulated regions where the central 1Mb had a recombination rate of less than $10^{-3}cM$ were excluded from the analysis.
    This excluded 7 regions from the analysis. 
    Then, power (or true positive rate) to detect a sweep by a given method was determined for
    100 evenly-spaced positions on chromosome 1 as follows:
    each sweep-detection method was applied to the 1,000 
    simulated datasets generated with a hard sweep (neutral and BGS) for that position,
    and the power was set to be the proportion of simulations (out of 1,000)
    in which the test statistic exceeded the threshold determined by each of the two null models (neutral and BGS).
    We note that using an empirical distribution to determine the threshold for calling a selective
    sweep with \diploshic deviates from the original implementation of the method,
    however for consistency among methods we chose to take this approach.
    The complete snakemake workflow for sweep detection can be found at
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/sweep_simulate.snake}.

\section*{Data availability}\label{data_availability}
The release of \stdpopsim is available at \url{https://github.com/popsim-consortium/stdpopsim}.
This paper coincides with version 0.3.XX of \stdpopsim. %be sure to check this
The entire analysis pipeline associated with this paper is available
as a snakemake workflow at \url{https://github.com/popsim-consortium/analysis2}.

>>>>

\newcommand{\tskit}{\texttt{tskit}\xspace}
\newcommand{\slim}{\texttt{SLiM}\xspace}
\newcommand{\msprime}{\texttt{msprime}\xspace}
\newcommand{\snakemake}{\texttt{snakemake}\xspace}

@@

    % current numbers of DFEs / species / etc
    % ADK: check version number
    At this release (version 0.3) the \stdpopsim catalog comprises 27 species (i.e. genome representations)
    for which we have included 35 demographic models, 38 genetic maps, 12 DFEs, and 8 annotation objects. %ADK: checked this on 3/10/25
    Additions to the catalog are ongoing---we point the reader to our previously published 
    report detailing this effort \citep{lauterbur2023expanding}. We welcome contributions from the
    community and encourage those interested in contributing to visit https://popsim-consortium.github.io/stdpopsim-docs
    or contact one of the PopSim Consortium members for guidance.

@@

    To do this we simulated human genomes with and without natural selection.
    In both scenarios, we ran three replicate simulations
    under the out of Africa (OOA) demographic model of
    \citep{ragsdale2019models} using a genetic map from the HapMap Project \citep{international2007second} (see \textbf{Methods}).
    Simulations with selection were implemented by modeling purifying selection on exons
    using a DFE inferred by \cite{kim2017inference}.
    A basic summary of nucleotide diversity from these simulations is shown in Figure \ref{fig:homsap-summary-stats},
    which shows that the presence of selection decreases polymorphism in simulations with selection only slightly
    at the genome-wide level, but more substantially within exonic regions under both a constant population size
    and the out-of-Africa demographic model.
    Following the simulation, inference of $N_e(t)$ was conducted from each of the six simulated datasets using four methods: 1) \msmc \citep{Schiffels2020}, 
    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
    % IG: I commented this out because it appeared distracting. We can add a note in the end of this analysis
    % However, as a thorough benchmarking is not the main purpose of this manuscript, we saved computational resources by not conducting more simulation replicates.

    Although the estimates produced by the four methods differ from one another, 
    for each method the estimates produced from data with selection (Figure
    \ref{fig:human-demography}; orange) and without selection (Figure
    \ref{fig:human-demography}; blue) are fairly similar.
    Thus, this particular parameterization of background selection
    in combination with selection only at exonic sites in the out-of-Africa demographic model does not
    appear to notably bias the estimates produced by these methods.
    In light of Figure \ref{fig:human-demography}, this makes sense-- 
    the effect of selection is only apparent within exonic regions and presumably sites
    very closely linked, thus masking of exons should be sufficient to remove the effects of selection
    from our simulations. 
    As noted in previous studies that compared demography inference methods
    on neutral simulations \citep{adrion2020community},
    we see that each method has its strengths and weaknesses.
    Methods based on the sequential Markovian coalescent (\smcpp and \msmc)
    appear overall to produce the most accurate trajectories of $N_e(t)$
    in this setting, although they tend to under-estimate the population
    bottleneck associated with the population split.
    On the other hand, \stairway, which utilizes the site frequency spectrum,
    produces more noisy estimates,
    but it does not appear to under-estimate the population bottleneck.
    The inference of \gone, which utilizes identity by descent,
    is targeted toward demographic changes in more recent time frames ($\sim 200$ generations),
    explaining its noisy estimates for all but the most recent time periods.

@@

    In comparing demography inference methods, we see that the SFS-based method (\stairway) produces more accurate
    inference than the two methods based on the sequential Markovian coalescent 
    (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
    Regardless of these differences, we see that background selection again does not
    appear to considerably influence the inference of any of the four methods for this species,
    at least for the parameterization used here, 
    even though vaquita has a very different evolutionary history than humans.

@@

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
        \caption{Inferred versus simulated DFEs from simulated human genomes.
        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
        acting on exons (see \textbf{Methods}).
        DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
        by three different methods: \grapes, \polydfe , and \dadicli.
        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
        shown for each DFE inferred from all three simulated datasets,
        with median values marked by horizontal bars
        and simulated values represented by dashed horizontal lines.
        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
        white bars represent the distribution used in the simulation.
        }
        \label{fig:homsap-dfe.ooa}
    \end{figure}

@@

    \begin{figure}[b!]
        \centering
        \includegraphics[width=0.8 \textwidth]{figures/sweeps/relationship_power_cM.pdf}
        \caption{
        Power to detect selective sweeps 
        %IG: shortened the title line
        %at 100 locations along human
        %chromosome 1 under a three population out-of-Africa model
        %\citep{gutenkunst2009inferring}  when using null models with and
        %without purifying selection
        as a function of local recombination rate.
        This figure shows the same power estimates shown in Figure \ref{fig:chr1_power},
        but with the 100 genomic segments sorted according to their
        average recombination rates instead of position along chromosome 1.
    	Genomic segments were simulated with sweeps under a three population out-of-Africa model
        \citep{gutenkunst2009inferring} and with background selection from deleterious mutations in exons.
        Three methods for detecting sweeps were applied to simulated data:
        1) \sweepfinder \citep{degiorgio2016sweepfinder2} (top row),
        2) \diploshic \citep{kern2018diplos} (middle row),
        and 3) reduced diversity ($\pi$) (bottom row).
        Power (true positive rate) is shown for these methods for the CEU and YRI
        samples (left and right respectively).
        The thresholds of the test statistics were set to control for
        $5\%$ false positive rate under a neutral null model (blue)
        and a null model with background selection from deleterious mutations in exons (red).
        Fitted lines represent loess smoothed regressions.
        }
        \label{fig:power-recomb}
    \end{figure}

@@

    The exploration of performance among methods done here is not meant to be thorough 
    benchmarking, but instead meant to highlight the utility of the \stdpopsim resource. 
    Having said that, there are a number of interesting results that bear discussion. 
    The first pertains to estimation of demographic history ($N_e(t)$) in the context of
    linked purifying selection. Our genomic simulations of humans and vaquita that incorporate
    selection model published DFEs that operate at exonic positions of the genome only. 
    As shown in Figures \ref{fig:homsap-summary-stats} and \ref{fig:vaquita-summary-stats},
    this combination only produces sizeable decreases in polymorphism at exonic positions, 
    which represent a small minority of sites within the genome, and does not lead to large-scale
    perturbations in genome-wide diversity. Thus it is not surprising that estimates of $N_e(t)$
    from genome-wide data from which exons have been masked should be accurate 
    (e.g., Figures \ref{fig:human-demography} and \ref{fig:vaquita-demography}). 
    However the question remains to what extent this would be true in empirical data where
    selection is not constrained to coding positions. For instance, we know
    that many genomes harbor large numbers
    of conserved non-coding elements \citep[e.g.,]{siepel2005evolutionarily} with extremely
    strong selection acting on them \citep{katzman2007human,mclean2008dispensability},
    thus we caution extrapolating from our simulation results. Future studies might build upon
    the annotations we used here to include selection at other portions of the genome as well
    as mixtures of DFEs across sites.

    A second important finding we made using the new stdpopsim selection API is that inference of the DFE
    is sensitive to the dominance of deleterious mutations.
    All three of the inference methods of the DFE applied here assume that deleterious mutations act
    in an additive fashion (h=0.5) by default.
    However, the simulated DFE for the vaquita genome included highly recessive mutations.
    Specifically, these simulations used a DFE where mutations with $s>0.01$ are quite recessive, 
    with $h<=0.01$, and where mutations with $0.001<s<0.01$ have $h=0.1$.
    We found that the three DFE inference methods underestimated $E[s]$ and overestimated the shape parameter (Figure \ref{fig:vaquita-dfe}).
    This performance is not surprising as decreasing $h$ for a given distribution of $s$ will lead to a decrease in $hs$,
    or an apparent decrease in the amount of selection.
    Fitting a mis-specified DFE that assumes $h=0.5$ will then lead to underestimates of $E[s]$.
    Indeed, the distribution of $hs$ is much more accurately estimated (Figure \ref{fig:vaquita-dfe}b).
    Overall, these results are consistent with prior work showing that it is challenging to separately
    infer the DFE of $s$ along with $h$ \citep{veeramah2014evidence, kyriazis2024constraining, balick2022overcoming}
    and that a large proportion ($>5\%$) of recessive deleterious mutations can confound inferences of $s$ \citep{wade2023quantifying}.
    Given the plethora of evidence for recessive deleterious mutations 
    \citep{mukai1972mutation, agrawal2011inferences, huber2018gene, di2024revisiting,},
    these results suggest the need for care in DFE inference.
    Kyriazis and Lohmueller \citeyear{kyriazis2024constraining} inferred the DFE of $s$ for nonsynonymous mutations
    in humans when considering different values of $h$.
    Intriguingly, additive models showed a satisfactory fit to the SFS of nonsynonymous variants in humans
    and more complex models of dominance did not substantially improve the fit.
    Nevertheless, models including recessive deleterious mutations, like those simulated here for the vaquita genome,
    show a similar fit to the SFS and may be more biologically realistic than fully additive models,
    given the evidence for the presence of recessive deleterious mutations.
    More broadly, new computational approaches are needed to precisely quantify the DFE and the covariance of $h$ and $s$.

    Further, inference of the DFE was done here using exon annotations of the genome in question.
    A key distinction between our analyses and those done in practice
    is that we have complete knowledge of which sites are under selection in the simulation,
    thus our inferences represent a best case scenario. Empirical inference of the DFE is a much
    harder problem as some mutations assumed to be free of selection
    may in reality be under selection.

    Regarding sweep detection, the inclusion of selection models into \stdpopsim allowed us to readily compare the
    effectiveness of different methods on data reflecting the demographic history, recombination rate variation,
    and genomic locations of negatively selected sites present in human populations. These results clearly demonstrate
    the impact of recombination rate variation, and underscore the importance of training methods on data with variable
    recombination rates \citep[e.g.,][]{schrider2017soft}. In addition, our results
    illustrate a strength of likelihood methods like \sweepfinder, that do not require a predefined window size.
    Window-based methods, such as \diploshic, tend to lag in performance when the perturbation due to a sweep is at 
    a much greater or smaller scale than the chosen window size, as can occur with very strong sweeps or low levels of 
    recombination. Moreover while all the methods examined here retain adequate power for sweep identification across
    scenarios, including the inclusion of background selection, our results underscore the importance of developing 
    methods that are well powered in the face of unknown demographic histories. Indeed, the largest differences
    in power we observed across simulated scenarios were between demographic histories and not the inclusion of background
    selection.

@@

    \subsection*{Whole genome simulations}
    In this study we performed simulations of the human and vaquita porpoise genomes
    using the \stdpopsim{} framework. We ran replicate datasets for each genome/parameter combination,
    in which chromosomes were simulated separately from one another.
    All simulations were executed using the \slim engine with a burn-in of $10N$ generations
    (where $N$ is the initial population size), and a scaling factor of 1 (no scaling).
    All simulations were implemented using as part of a \snakemake workflow,
    where the code utilized is similar to the one shown in Figure 1B.
    The complete simulation workflow is available at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/simulation.snake}
    (see \textbf{Data availablity} below).

    \subsubsection*{Simulations of complete human genomes}
    Simulation of human genomes used a genetic map from the HapMap Project
    \citep{international2007second} (\stdpopsim label \texttt{HapMapII\_GRCh38}) and
    the out of Africa (OOA) demographic model with archaic admixture from
    \cite{ragsdale2019models} (\stdpopsim label \texttt{OutOfAfricaArchaicAdmixture\_5R19}).
    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
    \cite{kim2017inference} (\stdpopsim label \texttt{Gamma\_K17}) to exons annotated by the HAVANA group release 104
    (\cite{ensembl2018}; \stdpopsim label \texttt{ensembl\_havana\_104\_exons}).
    Each simulation generated 100 diploid genomes from each of the three extant human populations (YRI, CEU, and CHB),
    with each genome consisting of 22 autosomes.
    Using these settings, we generated three replicate datasets without selection
    and three replicate datasets with selection.
    Three additional datasets, used in Figure \ref{fig:homsap-dfe.constant},
    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $10,000$.
    Each of these simulations generated 100 diploid genomes.
    The configuration file used in combination with our \snakemake workflow
    for these simulations can be found at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_HomSap.yml}.

    \subsubsection*{Simulations of complete vaquita porpoise genomes}
    Simulation of vaquita porpoise genomes used 
    the two-epoch demographic model from \cite{robinson2022critically} (\stdpopsim label \texttt{Vaquita2Epoch\_1R22}).
    Since no genetic map is currently available for this species, we assumed a constant recombination rate of  $10^{-8}$
    across all chromosomes \citep{morin2021}.
    Simulations with selection additionally applied a DFE model with gamma-distributed selection coefficients inferred by
    \cite{robinson2022critically} (\stdpopsim label \texttt{Gamma\_R22}) to exons annotated from the vaquita genome assembly
    (\cite{morin2021}; \stdpopsim label \texttt{Phocoena\_sinus.mPhoSin1.pri.110\_exons}).
    This DFE model implements a relationship between selection and dominance as follows:
    very deleterious mutations (with $s<-0.1$) were set to be recessive ($h=0$),
    mutations with $s\in [-0.1,-0.01)$ were set to have a dominance coefficient of $h=0.01$,
    mutations with $s\in [-0.01,-0.001)$ were set to have a dominance coefficient of $h=0.1$,
    and very mild mutations (with $s\in [-0.001,0)$) were set to have a nearly additive effect ($h=0.4$).
    This approach is similar to what was used in the simulations cunducted by \textcite{robinson2022critically}
    and is easily implemented in \stdpopsim's \texttt{DFE} class.
    Each simulation generate 100 diploid genomes,
    with each genome consisting of 21 autosomes.
    Using these settings, we generated three replicate datasets without selection
    and three replicate datasets with selection.
    Three additional datasets, used in Figure \ref{fig:vaquita-dfe.constant},
    were simulated with the same model of selection and a simple demographic model with one population with a constant size of $3,500$.
    \igcomment{If we end up using Fig. S3, then we should add a line describing it's simulation here.}
    The configuration file used in combination with our \snakemake workflow
    for these simulations can be found at 
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/config/snakemake/production_config_PhoSin.yml}.

@@

    We assessed the accuracy of demographic inferences from three software packages:
    \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent},
    and \smcpp \citep{terhorst2017robust}.
    For \msmc \citep{Schiffels2020}, we used a random sample of $N=6$ simulated genomes with 20
    iterations of the EM algorithm.
    For \gone \citep{santiago2020recent}, we set \texttt{max\_snps} to 500,000,
    the number of generations to 2,000, and the number of bins to 400, all representing default settings.
    For \stairway \citep{liu2020stairway} we did one run of theta estimation for each replicate dataset,
    and used a `dimFactor' of 5000, both of which represent default settings.
    For \smcpp \citep{terhorst2017robust} we used the unfolded SFS (`--unfold'), 
    along with the true mutation rate under which the simulations were generated. 
    The complete \snakemake workflow for demographic inference can be found at
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/n_t.snake}.

@@

    % ADK: I've added a section below that describes the DFE inference methods used in this study.
    For comparison among methods, estimates were rescaled. 
    As described above, each DFE-inference method scales selection coefficients slightly differently.
    We thus rescaled each program's output to obtain a selection coefficient ($s$) consistent with the one used by \slim.
    The output of \grapes  was scaled by a factor of  $\mu / \theta$,
    and the output of \polydfe was saled by $2 \mu / \theta$, 
    where $\mu$ is the assumed mutation rate and 
    $\theta$ is the population-scaled mutation rate estimated by each method ($\theta = 4N_{e}\mu$).
    For \dadicli, we report $\alpha * \nu / (\theta / (4 \mu L p))$, where
    $\alpha$ and $\nu$ are the inferred shape and rate parameters of the \dadicli's inferred gamma distribution,
    $L$ is the total number of sites available for the analysis,
    $p$ is the proportion of sites under selection. 
    Complete details for the DFE inference analysis can be found in the
    \snakemake workflow at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/dfe.snake}.

@@

    The three sweep-detection methods were associated with thresholds for their respective test statistics
    ($\pi$, \sweepfinder's CLR, or \diploshic's probability),
    which were determined using an empirical distribution representing a null model without sweeps.
    We considered two null models: a neutral evolution model (neutral),
    and a model with background selection from deleterious mutations in exons (BGS).
    Each null model was represented by a different set of $100,000$ simulations generated according to the appropriate model
    (see \textbf{Simulations with selective sweeps} above).
    Thus, for each test statistic we determined two thresholds corresponding to these two different null models (neutral and BGS)
    using the extreme $5\%$ quantile in each empirical distribution.
    Simulated regions where the central 1Mb had a recombination rate of less than $10^{-3}cM$ were excluded from the analysis.
    This excluded 7 regions from the analysis. 
    Then, power (or true positive rate) to detect a sweep by a given method was determined for
    100 evenly-spaced positions on chromosome 1 as follows:
    each sweep-detection method was applied to the 1,000 
    simulated datasets generated with a hard sweep (neutral and BGS) for that position,
    and the power was set to be the proportion of simulations (out of 1,000)
    in which the test statistic exceeded the threshold determined by each of the two null models (neutral and BGS).
    We note that using an empirical distribution to determine the threshold for calling a selective
    sweep with \diploshic deviates from the original implementation of the method,
    however for consistency among methods we chose to take this approach.
    The complete \snakemake workflow for sweep detection can be found at
    \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/sweep_simulate.snake}.

\section*{Data availability}\label{data_availability}
The release of \stdpopsim is available at \url{https://github.com/popsim-consortium/stdpopsim}.
This paper coincides with version 0.3.XX of \stdpopsim. %be sure to check this
The entire analysis pipeline associated with this paper is available
as a \snakemake workflow at \url{https://github.com/popsim-consortium/analysis2}.
